---
title: "Quiz3"
author: "Kiichi Takeuchi"
date: "September 19, 2014"
output:
  html_document:
    theme: united
    toc: yes
---


```{r,include=FALSE}
library(UsingR)
data(mtcars)
```

###Q1

Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as confounder. Give the adjusted estimate for the expected change in mpg comparing 8 cylinders to 4.

```{r}
#summary(lm(mpg~cyl+wt,data=mtcars))
fit<-lm(mpg~I(as.factor(cyl))+wt-1,data=mtcars)
fit
fit<-lm(mpg~I(as.factor(cyl))-1,data=mtcars)
fit
fit<-lm(mpg~
          I(1 * (cyl==4) )
         +I(1 * (cyl==8) )
         +wt
        ,data=mtcars)
fit
fit<-lm(mpg~
          I(1 * (cyl==4) )
         +I(1 * (cyl==8) )         
        ,data=mtcars)
fit

plot(mtcars[mtcars$cyl==4,]$wt,mtcars[mtcars$cyl==4,]$mpg,col="red",xlim=c(0,5),ylim=c(10,40))
fit1<-lm(mtcars[mtcars$cyl==4,]$mpg~mtcars[mtcars$cyl==4,]$wt)
abline(fit1,col="red")
fit1
points(mtcars[mtcars$cyl==8,]$wt,mtcars[mtcars$cyl==8,]$mpg,col="blue")
fit2<-lm(mtcars[mtcars$cyl==8,]$mpg~mtcars[mtcars$cyl==8,]$wt)
abline(fit2,col="blue")
fit2
#fit0<-lm(mpg~wt,data=mtcars)
#abline(fit0,col="green")
```

###Q2

Compare the adjusted by weight effect of 8 cylinders as compared to 4 the unadjusted. What can be said about the effect?.

```
With Wt: 8->4 diff is 6.07
Without Wt: 8->4 diff 11.56
```

###Q3

Fit a model with mpg as the outcome that considers number of cylinders as a factor variable and weight as confounder. Now fit a second model with mpg as the outcome model that considers the interaction between number of cylinders (as a factor variable) and weight. Give the P-value for the likelihood ratio test comparing the two models and suggest a model using 0.05 as a type I error rate significance benchmark.


```{r}
# as counfounder (as another varaiable)
fit1<-lm(mpg ~ factor(cyl) + wt, mtcars)
#same (expands like this)
#lm(mpg ~ factor(cyl) + wt + factor(cyl) : wt, mtcars)
summary(fit1)$coef

#star means interaction
fit2<-lm(mpg ~ factor(cyl) * wt, mtcars)
summary(fit2)$coef

anova(fit1,fit2)
``` 
Model #2 with interaction of weight is not significant; that is, we failed reject null hypthosis since p-value is 0.12 > 0.05.


 
###Q4

Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight inlcuded in the model as

```{r}
fit<-lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars)
summary(fit)$coef

plot(mtcars$wt,mtcars$mpg)
abline(fit,col="red")
abline(lm(mpg ~ I(wt * 0.6) + factor(cyl), data = mtcars),col="blue")
abline(lm(mpg ~ I(wt * 0.7) + factor(cyl), data = mtcars),col="green")
abline(lm(mpg ~ I(wt * 0.8) + factor(cyl), data = mtcars),col="lightblue")
abline(lm(mpg ~ I(wt * 0.9) + factor(cyl), data = mtcars),col="gray")
abline(lm(mpg ~ I(wt * 1.0) + factor(cyl), data = mtcars),col="purple")
```


###Q5

Give the hat diagonal for the most influential point

See Video 02_04_c Residuals and diagnostics examples 6:32

```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
fit<-lm(y~x)

hatvalues(fit)

plot(x,y,xlim=c(-1,13))
abline(fit,col="red")
text(x, y,sprintf("(%s,%s)",x,y), cex=0.6, pos=4, col="brown")
```



###Q6

Give the slope dfbeta for the point with the highest hat value.

Note: dfbetas shows values if you take out the data, how much the impact would be?

hatvalues is potential infulence within the dataset

See Video 02_04_c Residuals and diagnostics examples 6:32

```{r}
dfbetas(fit)
```


###Q7

Consider a regression relationship between Y and X with and without adjustment for a third variable Z. Which of the following is true about comparing the regression coefficient between Y and X with and without adjustment for Z.


```{r}
```


