---
title: "R Markdown Exercise"
author: "Kiichi Takeuchi"
date: "July 16, 2014"
output: html_document
---

This is my exercise to use R Markdown file with Knitr Package. 
The final product would be exported as HTML file.

```{r mycodechunk}
library(datasets)
data(airquality)
summary(airquality)
```

You should not see code chunk below:
```{r hiddenchunk,echo=FALSE,results='hide',cache=TRUE}
a<-rnorm(100)
# DO NOT SHOW THIS CHUNK
print(a)
# DO NOT EVEN SHOW THE RESULTS
```
Resuming R Markdown.


```{r}
rval<-rnorm(1)
```

Alias to random number from above: `r rval`

Trying pair function ...

```{r fig.height=10}
pairs(airquality)
```

Regression Model of Ozone over Solar Radiation, Wind and Temperature.

```{r}
library(stats)
fit <- lm(Ozone ~ Solar.R + Wind + Temp, data = airquality)
summary(fit)
```

Here, use xtable to format the results

```{r results='asis'}
# Install xtable package
library(xtable)
xt <- xtable(summary(fit))
print(xt,type='html')
```

#Summary

* Looks like Ozone is correlated with Temperature.
* And so on.

#How to set code chunk Options Globally?

```{r}
# somehow thid does not work below??
# opts_chunk$set(echo=FALSE,results="hide")
```

No more verbosing results below...
```{r}
print(100)
```








