```{r}
library(datasets)
summary(mtcars$mpg)
#hist(mtcars$mpg)
t.test(mtcars$mpg)
t.test(mtcars$mpg)$conf.int

#If you need the t interval of a specific confidence interval, e.g. 97.5%
t.test(mtcars$mpg,conf.level=0.975)
```

Suppose that standard deviation of 9 paired differences is 1, what value would the average difference have to be so that the lower endpoint of a 95% students t confidence interval touch zero?
```{r}
# See Gosset's t distr slide
#degree of freedom n-1: 8
#S is 1
qt(.975, df = 8) * 1 / sqrt(9)
```

Slide6:Consider the mtcars dataset. Construct a 95% T interval for MPG comparing 4 to 6 cylinder cars (subtracting in the order of 4 - 6) assume a constant variance.
```{r}
four<-mtcars[mtcars$cyl==4,]$mpg
six<-mtcars[mtcars$cyl==6,]$mpg
t.test(four,six,var.equal = T) #assume a constant variance

```
The interval is above zero, suggesting 4 is better than 6 in the terms of MPG.
If you switch four and six (six - four) the t interval goes negative.


Suppose that 18 obese subjects were randomized, 9 each, to a new diet pill and a placebo. Subjects' body mass indices (BMIs) were measured at a baseline and again after having received the treatment or placebo for four weeks. The average difference from follow-up to the baseline (followup - baseline) was 3 kg/m2 for the treated group and 1 kg/m2 for the placebo group. The corresponding standard deviations of the differences was 1.5 kg/m2 for the treatment group and 1.8 kg/m2 for the placebo group. The study aims to answer whether the change in BMI over the four week period appear to differ between the treated and placebo groups.
```{r}
n1 <- n2 <- 9
x1 <- -3  ##treated
x2 <- 1  ##placebo
s1 <- 1.5  ##treated
s2 <- 1.8  ##placebo
( (n1 - 1) * s1^2 + (n2 - 1) * s2^2) / (n1 + n2 - 2)
```