---
title: "quiz3"
author: "Kiichi Takeuchi"
date: "October 25, 2014"
output:
  html_document:
    theme: united
    toc: yes
  pdf_document:
    toc: yes
---


###Q1
1. Subset the data to a training set and testing set based on the Case variable in the data set. 
2. Set the seed to 125 and fit a CART model with the rpart method using all predictor variables and default caret settings. 
3. In the final model what would be the final model prediction for cases with the following variable values:
a. TotalIntench2 = 23,000; FiberWidthCh1 = 10; PerimStatusCh1=2 
b. TotalIntench2 = 50,000; FiberWidthCh1 = 10;VarIntenCh4 = 100 
c. TotalIntench2 = 57,000; FiberWidthCh1 = 8;VarIntenCh4 = 100 
d. FiberWidthCh1 = 8;VarIntenCh4 = 100; PerimStatusCh1=2 

```{r}
library(AppliedPredictiveModeling)
data(segmentationOriginal)
library(caret)

summary(segmentationOriginal$Case)

training = segmentationOriginal[segmentationOriginal$Case=='Train',]
testing = segmentationOriginal[segmentationOriginal$Case=='Test',]

set.seed(125)

fit<-train(Species~.,method="rpart",data=training)
```



###Q2



###Q3

These data contain information on 572 different Italian olive oils from multiple regions in Italy. Fit a classification tree where Area is the outcome variable. Then predict the value of area for the following data frame using the tree command with all defaults

```{r}
library(pgmm)
setwd("~/work/r/class/datasciencecoursera/notes/class8_predmachlearn/week3")
load('olive.rda')
data(olive)
olive = olive[,-1]
#Note, if you don't apply factor, it will calculate the area as Numeric!
#olive$Area = factor(olive$Area)
head(olive)

newdata = as.data.frame(t(colMeans(olive)))
newdata

inTrain = createDataPartition(olive$Area, p = 0.7, list=F)
training = olive[inTrain,]
testing = olive[-inTrain,]
qplot(Linoleic,Linolenic,col=Area,data=training)
fit<-train(Area~.,method="rpart",data=training)
fit$finalModel
library(rattle)
fancyRpartPlot(fit$finalModel)

predict(fit,newdata)
````
What is the resulting prediction? Is the resulting prediction strange? Why or why not?

###Q4


Then set the seed to 13234 and fit a logistic regression model (method="glm", be sure to specify family="binomial") with Coronary Heart Disease (chd) as the outcome and age at onset, current alcohol consumption, obesity levels, cumulative tabacco, type-A behavior, and low density lipoprotein cholesterol as predictors. Calculate the misclassification rate for your model using this function and a prediction on the "response" scale:

What is the misclassification rate on the training set? What is the misclassification rate on the test set?
```{r}
library(ElemStatLearn)
data(SAheart)
set.seed(8484)
#SAheart$chd<-factor(SAheart$chd)
intrain = sample(1:dim(SAheart)[1],size=dim(SAheart)[1]/2,replace=F)
trainSA = SAheart[intrain,]
testSA = SAheart[-intrain,]

library(caret)
set.seed(13234)

fit<-train(chd~.,method="glm",family="binomial",data=trainSA)
fit
#est<-predict(fit,testSA)
#c(length(testSA$chd),length(prediction))
#confusionMatrix(testSA$chd,prediction)
missClass = function(values,prediction){
  sum(((prediction > 0.5)*1) != values)/length(values)
}
missClass(testSA$chd,predict(fit,testSA))
missClass(trainSA$chd,predict(fit,trainSA))

```


###Q5



